---
title: "Econometrics Lab 6"
format: pdf
editor: visual
---

| Header 1 | Header 2 | Header 3 |
|----------|----------|----------|
| 1        | 2        | 3        |

: Demonstration of Markdown Table Syntax

| Header 1 | Header 2 | Header 3 | Header 4 |
|---------:|:---------|----------|:--------:|
|    Right | Left     | Default  |  Center  |

: Demonstration of Markdown table syntax with aligned columns

| Header 1 | Header 2 | Header 3 |
|----------|----------|----------|
| 1        | 2        | 3        |

: Table for reference {#table-example}

+---------------+--------------------+
| # Col1        | # Col2             |
+---------------+--------------------+
| 12            |                    |
+---------------+--------------------+
| 5462          | 77                 |
+---------------+--------------------+
|               | 99                 |
+---------------+--------------------+

: Table using the Visual Editor

\begin{tabular}{lcr}
Header 1 & Header 2 & Header 3\\
Left & Center & Right\\
\end{tabular}


\begin{tabular}{l|c|r}
\hline
Header 1 & Header 2 & Header 3\\
\hline
Left & Center & Right\\
\hline
\end{tabular}
\end{center}


\begin{table}[!h]
\caption{This is a caption}
\begin{center}
\begin{tabular}{l|c|r}
\hline
Header 1 & Header 2 & Header 3\\
\hline
Left & Center & Right\\
\hline
\end{tabular}
\end{center}
\end{table}

```{r}

  
d <- data.frame("Header 1" = c("Left",1),
                "Header 2" = c("Center",2),
                "Header 3" = c("Right", 3),
                check.names = FALSE)
knitr::kable(d,
             format = "pipe",
             booktabs = T,
             escape = F,
             align = c("l","c","r"),
             longtable = T,
             caption = "kable Table")
```

```{r}

library(kableExtra)

d <- data.frame("Header 1" = c("Left",1),
                "Header 2" = c("Center",2),
                "Header 3" = c("Right", 3),
                check.names = FALSE)
knitr::kable(d,
             format = "latex",
             booktabs = T,
             escape = F,
             align = c("l","c","r"),
             longtable = T,
             caption = "kable Table Centered")%>%
  kable_styling(position = "center")
```

